---
title: 'Reproducible Research: Peer Assessment 1'
author: "Antonio Guerreiro"
date: "29/07/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
activity$day_interval <- activity$interval %% 100
nona_activity = na.omit(activity)
```

## What is mean total number of steps taken per day?

```{r}
aggregate(nona_activity$steps, list(nona_activity$date), sum)
aggregate(nona_activity$steps, list(nona_activity$date), mean)
aggregate(nona_activity$steps, list(nona_activity$date), median)
hist(aggregate(nona_activity$steps, list(nona_activity$date), sum)$x)
```

Total mean and median

mean

```{r}
mean(nona_activity$steps)
```

median

```{r}
median(nona_activity$steps)
```

## What is the average daily activity pattern?

```{r}
nona_activity$day_interval <- nona_activity$interval %% 100
plot(aggregate(nona_activity$steps, list(nona_activity$day_interval), mean)$Group.1, aggregate(nona_activity$steps, list(nona_activity$day_interval), mean)$x, type = "l")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
interval_means <-aggregate(nona_activity$steps, list(nona_activity$day_interval), mean)
interval_means
interval_means[interval_means$x == max(interval_means$x),]
```


## Imputing missing values

Total number of rows with  NAs

```{r}
NROW(which (is.na(activity)))
nrow(activity[is.na(activity$steps),])
```

The strategy is to use the the mean for that 5-minute interval.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
for (n in 1:nrow(activity)) {
  if (is.na(activity$steps[n])) {
    new_value <- interval_means[interval_means$Group.1 == activity$day_interval[n], ]$x
    activity$steps[n] <- new_value
  }
}

head(activity)
```

Total number of rows with  NAs

```{r}
NROW(which (is.na(activity)))
nrow(activity[is.na(activity$steps),])
```

```{r}
aggregate(activity$steps, list(activity$date), mean)
aggregate(activity$steps, list(activity$date), median)
hist(aggregate(activity$steps, list(activity$date), sum)$x)
```

Do these values differ from the estimates from the first part of the assignment? yes

What is the impact of imputing missing data on the estimates of the total daily number of steps?


total number of steps taken each day

```{r}
aggregate(activity$steps, list(activity$date), sum)
```


```{r}
aggregate(activity$steps, list(activity$date), sum)
aggregate(activity$steps, list(activity$date), mean)
aggregate(activity$steps, list(activity$date), median)
hist(aggregate(activity$steps, list(activity$date), sum)$x)
```

Total mean and median

mean

```{r}
mean(activity$steps)
```

median

```{r}
median(activity$steps)
```

What is the impact of imputing missing data on the estimates of the total daily number of steps?
No impact for overallmean and median.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity$day <- weekdays(as.Date(activity$date),abbreviate = FALSE)

for (n in 1:nrow(activity)) {
  if (activity$day[n] == "Samedi" || activity$day[n] == "Dimanche") {
    activity$daytype[n] <- "weekend"
  } else {
    activity$daytype[n] <- "weekday"
  }
}

head(activity)
```

panel plot

```{r}
weekend_activity <- activity[activity$daytype == "weekend",]
weekday_activity <- activity[activity$daytype == "weekday",]

head(weekend_activity)
head(weekday_activity)
```


```{r}

par(mfrow=c(2,1))
plot(aggregate(weekend_activity$steps, list(weekend_activity$interval), mean)$Group.1, aggregate(weekend_activity$steps, list(weekend_activity$interval), mean)$x, type = "l")

plot(aggregate(weekday_activity$steps, list(weekday_activity$interval), mean)$Group.1, aggregate(weekday_activity$steps, list(weekday_activity$interval), mean)$x, type = "l")


```


